---
title: "Mawikwe/Siskiwit Site Comparison"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(targets)
library(tidyverse)
library(plotly)
library(DT)
library(leaflet)

dnr <- tar_read(dnr, store = "../_targets") %>% 
  select(-c(nh3, doc, do,  chl_field, depth, source, type)) %>% 
               filter(str_detect(site, "site")) %>% 
               mutate(site = str_split_i(site, " ", 2),
                      site = as.factor(site)) %>% 
  filter(site %in% c("14", "15"))









StatChull <- ggproto("StatChull", Stat,
  compute_group = function(data, scales) {
    data[chull(data$x, data$y), , drop = FALSE]
  },
  
  required_aes = c("x", "y")
)
stat_chull <- function(mapping = NULL, data = NULL, geom = "polygon",
                       position = "identity", na.rm = FALSE, show.legend = NA, 
                       inherit.aes = TRUE, ...) {
  layer(
    stat = StatChull, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

.get_withinSS <- function(d, cluster){
  d <- as.dist(d)
  cn <- max(cluster)
  clusterf <- as.factor(cluster)
  clusterl <- levels(clusterf)
  cnn <- length(clusterl)
  
  cwn <- cn
  # Compute total within sum of square
  dmat <- as.matrix(d)
  within.cluster.ss <- 0
  for (i in 1:cn) {
    cluster.size <- sum(cluster == i)
    di <- as.dist(dmat[cluster == i, cluster == i])
    within.cluster.ss <- within.cluster.ss + sum(di^2)/cluster.size
  }
  within.cluster.ss
}

opt_clus <- function(x, kmax){
  if (is.data.frame(x)) x <- as.matrix(x)
  diss <- dist(x)
  
  v <- rep(0, kmax)
  
  for(i in 1:kmax){
    clust <- kmeans(x, i, nstart = 100)
    v[i] <- .get_withinSS(diss, clust$cluster)
  }
  
  df <- data.frame(clusters = as.factor(1:kmax), y = v, stringsAsFactors = TRUE)
  
  ggplot(data = df, aes(x = clusters, y = y, group = 1)) +
    geom_point() + 
    geom_line() +
    xlab("Number of clusters k") +
    ylab("Total Within Sum of Square") +
    ggtitle("Optimal number of clusters") +
    theme_classic()
  
}
# Function to calculate Euclidean distance
euclidean_distance <- function(df1, df2) {
  sqrt(sum((df1 - df2) ^ 2))
}

# Function to evaluate subsets
evaluate_subsets <- function(data, subset_size) {
  n <- nrow(data)
  all_subsets <- combn(n, subset_size, simplify = FALSE)
  
  # Calculate the mean of the entire data frame
  overall_mean <- colMeans(data)
  
  # Initialize a list to store results
  results <- list()
  
  for (subset_indices in all_subsets) {
    subset_data <- data[subset_indices, ]
    subset_mean <- colMeans(subset_data)
    
    # Calculate the distance from the overall mean
    distance <- euclidean_distance(overall_mean, subset_mean)
    
    # Store the indices as a single string and the distance in the results list
    results[[length(results) + 1]] <- data.frame(
      indices = paste(subset_indices, collapse = ", "),
      distance = distance
    )
  }
  
  # Combine all results into a single data frame
  results_df <- do.call(rbind, results)
  return(results_df)
}
```

## Key Takeaways
- Chlorophyll and Turbidity are by far the most variable between sites (42% of days had notable differences between sites), followed by TP (26%).
- There were far fewer cases where the trend from one date to another differed bewteen sites.
- All other parameters had few differences between sites.
- This implies that chlorophyll concentrations in the two bays is dependent somewhat on P loading as well as additional parameters not measured (light, warming, tributary seeding, etc.).
- Sampling both sites could be useful if we are able to better measure the local differences in some of those additional parameters.
- Given that the trends between sites are almost always the same, I think that we gain more from more frequent sampling than from sampling both, given the choice. 
- Running this analysis reminded me that the Siskiwit river does have a couple USGS gages, which could make sampling site 14 more valuable than 15, depending on how much of the water from the river makes it to Mawikwe Bay.

## Trends
### Chlorophyll Over time
```{r chl time}
dnr_range <- dnr %>% 
  select(date, chl) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)

ggplotly(
  ggplot(data = dnr, aes(x = date, y = chl, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on `r dnr_range$date`. (`r perc_days`% of days where both sites were sampled). Only one of those dates (2022-08-23) had a different trend (up/down vs previous sample) between sites. 

### TP Over time
```{r tp time}
dnr_range <- dnr %>% 
  select(date, tp) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)

ggplotly(
  ggplot(data = dnr, aes(x = date, y = tp, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on `r dnr_range$date`. (`r perc_days`% of days where both sites were sampled). Two of those dates (2022-08-23, 2023-07-31) had a different trend (up/down vs previous sample) between sites. 


### TN Over time
```{r tn time}
dnr_range <- dnr %>% 
  select(date, tn) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)


ggplotly(
  ggplot(data = dnr, aes(x = date, y = tn, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on `r dnr_range$date`. (`r perc_days`% of days where both sites were sampled).

### SRP Over time
```{r po4 time}
dnr_range <- dnr %>% 
  select(date, po4) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)


ggplotly(
  ggplot(data = dnr, aes(x = date, y = po4, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on `r dnr_range$date`. (`r perc_days`% of days where both sites were sampled)

### NO3 Over time
```{r no3 time}
dnr_range <- dnr %>% 
  select(date, no3) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)


ggplotly(
  ggplot(data = dnr, aes(x = date, y = no3, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on no days.



### TSS Over time
```{r tss time}
dnr_range <- dnr %>% 
  select(date, tss) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)


ggplotly(
  ggplot(data = dnr, aes(x = date, y = tss, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on `r dnr_range$date`. (`r perc_days`% of days where both sites were sampled) However, both of these days would have been non-detects in previous years.

### Turbidity Over time
```{r turb time}
dnr_range <- dnr %>% 
  select(date, turb) %>% 
  group_by(date) %>%
  na.omit() %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  select(-n) %>% 
  summarise(across(everything(), list(mean = mean, range = ~(max(.)-min(.))), .names = "{.fn}")) %>% 
  mutate(perc = range / mean) 

mult_count <- nrow(dnr_range)

dnr_range <- dnr_range %>% 
  filter(perc > .2)

perc_days <- round(nrow(dnr_range)/mult_count * 100)


ggplotly(
  ggplot(data = dnr, aes(x = date, y = turb, color = site)) +
    geom_point() +
    geom_line()
)
```

Notable differences (range > 20% of mean) between sites were present on `r dnr_range$date`. (`r perc_days`% of days where both sites were sampled)

There were no notable differences between sites for temperature, DO, conductivity, or ph.
