---
title: "CB Data Summary"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(targets)
library(tidyverse)
library(plotly)
library(DT)
library(gganimate)
library(sf)
library(Hmisc)
library(corrplot)
library(randomForest)
library(randomForestExplainer)

ls_crop <- st_crop(tar_read(ls_shp, store = "../_targets"), st_bbox(c(xmin = -91.1, ymin = 46.4, xmax = -90.6, ymax = 46.9)))
lake <- tar_read(cbnut_clean, store = "../_targets")

lake_corr <- select(lake, -c(date, site, source, type)) %>% 
      st_drop_geometry()
lake_cor <- rcorr(as.matrix(lake_corr), type = 'spearman')
lake_cor$P[is.na(lake_cor$P)] <- 1

lake_rf_prep1 <- lake %>% 
  st_drop_geometry() %>%
  mutate(month = month(date)) %>% 
  select(month, tp, po4, tss, temp, do_sat, ph, cond, chl) %>% 
  mutate(month = as.factor(month)) %>% 
  na.omit()
lake_rf1 <- randomForest(chl ~ ., data = lake_rf_prep1, importance = TRUE, ntree = 1000)
lake_rf_prep2 <- lake %>% 
  st_drop_geometry() %>%
  mutate(month = month(date)) %>% 
  select(month, pp, tdp, nh3, no3, tp, po4, tss, temp, do_sat, ph, cond, chl) %>% 
  mutate(month = as.factor(month)) %>% 
  na.omit()
lake_rf2 <- randomForest(chl ~ ., data = lake_rf_prep2, importance = TRUE, ntree = 1000)

```


## Correlations between variables

Here's a plot showing spearman correlation coefficients between variables. Bigger size/darker color means a stronger correlation. Only correlations with P < 0.05 are shown.

#### Lake Superior
```{r lake_corr, echo=FALSE, error=TRUE}
corrplot(lake_cor$r, type = "lower", tl.cex = .5, p.mat = lake_cor$P, sig.level = 0.05, insig = 'blank', outline = T)
```



## Variable Importance (random forest)
#### more limited parameters, more observations (760)
```{r rf_lake1, echo=FALSE}
# importance <- measure_importance(lake_rf1) %>%
#   mutate(mse_increase = round(mse_increase, 3),
#          mean_min_depth = round(mean_min_depth, 2),
#          node_purity_increase = round(node_purity_increase, 2)) %>%
#   select(-c(p_value, no_of_trees))
datatable(lake_rf1$importance)
```

#### more parameters, fewer observations (130)
```{r rf_lake2, echo=FALSE}
# importance <- measure_importance(lake_rf2) %>%
#   mutate(mse_increase = round(mse_increase, 3),
#          mean_min_depth = round(mean_min_depth, 2),
#          node_purity_increase = round(node_purity_increase, 2)) %>%
#   select(-c(p_value, no_of_trees))
datatable(lake_rf2$importance)
```

## Plots of Chlorophyll vs everything else
#### In CB
```{r chlplots, echo=FALSE}

ggplotly(lake %>%
  ggplot(aes(x = no3, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = nh3, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = pon, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = po4, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = tdp, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = pp, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = tp, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = poc, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = tss, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = temp, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = do_sat, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = ph, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = cond, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = turb, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
      mutate(month = as_factor(month(date))) %>%      
  ggplot(aes(x = month, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
```


## Maps!

```{r mapchl, echo=FALSE, fig.height=6}
plotchl <-  ggplot() +
               theme_bw() +
               geom_sf(data = ls_crop) +
               geom_sf(data = filter(lake, !is.na(chl)) %>% mutate(chl = if_else(chl > 20, NA, chl)), mapping = aes(color = chl), size = 7) +
               scale_color_gradient(low = "#d7f4cb", high = "darkgreen", na.value = "gray30") +
               transition_states(date) +
               labs(title = "CHL on Week of {closest_state}")
animate(plotchl, nframes = length(unique(lake$date)), fps = 2)
```