---
title: "NPS Buoy Summary"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(targets)
library(tidyverse)
library(plotly)
library(DT)
library(gganimate)
library(sf)
library(Hmisc)
library(corrplot)
library(randomForest)
library(randomForestExplainer)

#ls_crop <- tar_read(ls_crop, store = "../_targets")
lake <- tar_read(nps_sonde_wetlab, store = "../_targets") %>% 
  ungroup()

lake_trim <- select(lake, -c(date, site, depth, latitude, longitude)) 
lake_cor <- rcorr(as.matrix(lake_trim), type = 'spearman')
lake_cor$P[is.na(lake_cor$P)] <- 1

lake_rf_prep <- lake %>%
  mutate(month = month(date)) %>%
  select(-c(date, site, depth, latitude, longitude, tds)) %>%
  mutate(month = as.factor(month)) %>%
  na.omit()
lake_rf <- randomForest(chl ~ ., data = lake_rf_prep, importance = TRUE, ntree = 100)

```


## Correlations between variables

Here's a plot showing spearman correlation coefficients between variables. Bigger size/darker color means a stronger correlation. Only correlations with P < 0.05 are shown.

#### Lake Superior
```{r lake_corr, echo=FALSE, error=TRUE}
corrplot(lake_cor$r, type = "lower", tl.cex = .5, p.mat = lake_cor$P, sig.level = 0.05, insig = 'blank', outline = T)
```



## Variable Importance (random forest)
#### Lake Superior 
```{r rf_lake, echo=FALSE}
# importance <- measure_importance(lake_rf) %>%
#   mutate(mse_increase = round(mse_increase, 3),
#          mean_min_depth = round(mean_min_depth, 2),
#          node_purity_increase = round(node_purity_increase, 2)) %>%
#   select(-c(p_value, no_of_trees))
datatable(lake_rf$importance)
```

## Plots of Chlorophyll vs everything else
#### In Lake Superior
```{r chlplots, echo=FALSE}
ggplotly(lake %>%
  ggplot(aes(x = turb, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = cond, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = ph, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = do_sat, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = do, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = temp, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
  ggplot(aes(x = cdom, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
ggplotly(lake %>%
      mutate(month = as_factor(month(date))) %>%      
  ggplot(aes(x = month, y = chl, text = str_c(site, date, sep = " "))) +
  geom_point())
```


## Maps!

```{r mapchl, echo=FALSE, fig.height=6}
# lakechl <- filter(lake, !is.na(chl)) %>% mutate(chl = if_else(chl > 20, NA, chl))
# plotchl <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = lakechl, mapping = aes(color = chl), size = 7) +
#                scale_color_gradient(low = "#d7f4cb", high = "darkgreen", na.value = "gray30") +
#                transition_states(date) +
#                labs(title = "CHL on {closest_state}")
# animate(plotchl, nframes = length(unique(lakechl$date)), fps = 2)
```
```{r maptss, echo=FALSE, fig.height=2}
# plottss <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(umd, !is.na(tss)), mapping = aes(color = tss), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "tan4") +
#                transition_states(date) +
#                labs(title = "TSS on Week of {closest_state}")
# animate(plottss, nframes = length(unique(lake$date)), fps = 2)
```
```{r maptp, echo=FALSE, fig.height=2}
# plottp <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(tp)), mapping = aes(color = tp), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "darkblue") +
#                transition_states(date) +
#                labs(title = "TP on Week of {closest_state}")
# animate(plottp, nframes = length(unique(lake$date)), fps = 1)
```
```{r maptn, echo=FALSE, fig.height=2}
# plottn <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(tn)), mapping = aes(color = tn), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "darkblue") +
#                transition_states(date) +
#                labs(title = "TN on Week of {closest_state}")
# animate(plottn, nframes = length(unique(lake$date)), fps = 1)
```
```{r mapno3, echo=FALSE, fig.height=2}
# plotno3 <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(no3)), mapping = aes(color = no3), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "darkblue") +
#                transition_states(date) +
#                labs(title = "NO3 on Week of {closest_state}")
# animate(plotno3, nframes = length(unique(lake$date)), fps = 1)
```
```{r mapturb, echo=FALSE, fig.height=2}
# plotturb <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(turb)), mapping = aes(color = turb), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "tan4") +
#                transition_states(date) +
#                labs(title = "Turbidity on Week of {closest_state}")
# animate(plotturb, nframes = length(unique(lake$date)), fps = 1)
```
```{r mapcond, echo=FALSE, fig.height=2}
# plotcond <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(cond)), mapping = aes(color = cond), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "darkblue") +
#                transition_states(date) +
#                labs(title = "Conductivity on Week of {closest_state}")
# animate(plotcond, nframes = length(unique(lake$date)), fps = 1)
```
```{r mapph, echo=FALSE, fig.height=2}
# plotph <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(ph)), mapping = aes(color = ph), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "darkblue") +
#                transition_states(date) +
#                labs(title = "pH on Week of {closest_state}")
# animate(plotph, nframes = length(unique(lake$date)), fps = 1)
```
```{r mapdo, echo=FALSE, fig.height=2}
# plotdo <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(do)), mapping = aes(color = do), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "darkblue") +
#                transition_states(date) +
#                labs(title = "DO on Week of {closest_state}")
# animate(plotdo, nframes = length(unique(lake$date)), fps = 1)
```
```{r maptemp, echo=FALSE, fig.height=2}
# plottemp <-  ggplot() +
#                theme_bw() +
#                geom_sf(data = ls_crop) +
#                geom_sf(data = filter(lake, !is.na(temp)), mapping = aes(color = temp), size = 7) +
#                scale_color_gradient(low = "lightblue", high = "indianred2") +
#                transition_states(date) +
#                labs(title = "Temperature on Week of {closest_state}")
# animate(plottemp, nframes = length(unique(lake$date)), fps = 1)
```

